\texttt EPQ Key Exchange

### 2017-04-26 latex/keyexchange/secretkeybasedpublickeyencryption.tex
backing up work

\documentclass[preprint]{iacrtrans}
\usepackage[utf8]{inputenc}

@@

\author{Ella Rose\inst{1}}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[\texttt Lattice Based Key Exchange]{\texttt Lattice Based Key Exchange}

@@

\begin{abstract}
  We define an algorithm for performing a key exchange between two parties utilizing lattice based ideas. The algorithm makes use of a small number of bignum multiplications and additions and is conceptually simple.\\ 
\end{abstract}

@@

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance to the function of the internet; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.\\

@@

Lattices are also very attractive because they offer homomorphic properties, which is another of the most important topics in cryptography at the moment. This property is also extremely important because it allows for the construction of simple public key encryption schemes and key exchange shemes from secret key ciphers. We present such a scheme in this paper, as well as a homomorphic cipher suitable for instantiating the scheme.

\section{Definitions}
We use $+$ to denote addition, $ab$ to denote multiplication, and $\mod p$ to denote the modulo operation.

\section{Algorithm}
First, we shall define the key exchange algorithm; Then, we shall define a secret key cipher suitable for instantiating the key exchange algorithm.

@@

Where $r_1$ and $r_2$ are randomly generated values of appropriate size.\todo{What sizes are appropriate?} Then the value $c$ is shared with the owner of the public key. The owner of the public key can recover $e$ by applying the homomorphic decryption operation of the secret key cipher to c; This is because $p_1$ is an encryption of $0$, $p_1 r_1$ is equivalent to $0 r_1 \equiv 0$, similar for $p_2$. So when the decryption operation is applied, the result is $0 + 0 + e \equiv e$.

\subsection{Secret Key Cipher}
Creation of the $p_1$ and $p_2$ values requires a secure, additively homomorphic cipher. We define a simple lattice based cipher which utilizes effectively the same operations as the key exchange process. Let $s_1$ and $s_2$ be two large numbers, where $log_2(s_1) >= log_2(s_2) * 3$; These two numbers constitute the secret key for the cipher, which in turn acts as the private key for the key exchange presented previously.\\

@@

\subsection{Note}
While the decryption circuit works via two modulo operations of the base points of the lattice, this operation does not work for the public key/key exchange portion of the algorithm - The reason is because in the secret key cipher, the two different lattice points are of largely different size; In the key exchange algorithm, both points are roughly the same size, and so the double modulo operation does not function identically.

>>>>

\documentclass[preprint]{iacrtrans}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}

@@

\author{Ella Rose\inst{1}}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[\texttt EPQ Key Exchange]{\texttt EPQ Key Exchange}

@@

\begin{abstract}
  We define an algorithm for performing a key exchange between two parties. The algorithm makes use of a small number of bignum multiplications and additions and is conceptually simple.\\ 
\end{abstract}

@@

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance to the function of the internet; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.

@@

Lattices are also very attractive because they offer homomorphic properties, which is another of the most important topics in cryptography at the moment. This property is important because it allows for the construction of simple public key encryption schemes and key exchange shemes from secret key ciphers. We present such a scheme in this paper, as well as a homomorphic cipher suitable for instantiating the scheme.

\section{Definitions}
We use $+$ to denote integer addition, $ab$ to denote integer multiplication, and $\mod p$ to denote the modulo operation. We use $log(x)$ to indicate the binary logarithm of $x$.

\section{Algorithm}
First, we shall define the asymmetric key exchange algorithm; Then, we shall define a symmetric secret key cipher suitable for instantiating the key exchange algorithm.

@@

Where $r_1$ and $r_2$ are randomly generated values of appropriate size. Then the value $c$ is shared with the owner of the public key. The owner of the public key can recover $e$ from $c$ by applying the homomorphic decryption operation of the secret key cipher to c. Because $p_1$ is an encryption of $0$, $p_1 r_1 \equiv 0$, similar for $p_2$. So when the decryption operation is applied, the result is $0 + 0 + e \equiv e$.

\subsubsection{Additional Information for Key Generation}
If the size of the shared secret is to be 256 bits, then the ciphertexts produced by symmetric cipher must support the homomorphic addition of a plaintext 256-bit value.

Not all homomorphic constructions support the addition of plaintext values to ciphertext values; Lattices offer this required feature.

It is important that $p_1$ and $p_2$ are coprime to each other. Otherwise, some bits of the secret may be recovered by reducing a ciphertext modulo the greatest common divisor. Specifically:

\begin{align}
    e \gets secret\\
    c \gets (p_1 r_1) + (p_2 r_2) + e\\
    e_0 \gets c \mod gcd(p_1, p_2)
\end{align}

Where $log(e_0) \equiv log(gcd(p_1, p_2))$.

\subsection{Secret Key Cipher}
Creation of the $p_1$ and $p_2$ values requires a secure, additively homomorphic cipher. We define a simple lattice based cipher which utilizes effectively the same operations as the key exchange process. Let $s_1$ and $s_2$ be two large numbers, where $log(s_1) >= log(s_2) * 3$; These two numbers constitute the secret key for the cipher, which in turn acts as the private key for the key exchange presented previously.\\

@@

\subsubsection{Additional Information for Key Generation}
$s_1$ and $s_2$ should be coprime, and $log(s_1) >= log(s_2) * 3$

\subsection{Note}
While the decryption circuit works via two modulo operations of the base lattice points, this operation does not work for the public key/key exchange portion of the algorithm - The reason is because in the secret key cipher, the two different lattice points are of largely different size; In the key exchange algorithm, both points are roughly the same size, and so the double modulo operation does not function identically.

\subsection{Public Key Randomization}
Because the public key consists of malleable homomorphic ciphertexts, it can be randomized without complicating any of the rest of the procedures. The presented scheme can utilize encrypted public keys, and we give two potential use cases for the idea.

First, as an additional layer of security, users who wish to send a message to the owner of the public key can first randomize the public key that they are using. Assuming that the randomized $p_1$ and $p_2$ are in effect unrelated to the original $p_1$ and $p_2$, then observers of the ciphertext do not even have the base points that the ciphertext was created with.

Secondly, this feature allows for the disassociation of the owner of the public key from the possessor of the public key. For example, assume we have three parties, Alice, Bob, and Charlie. Charlie does not want Alice to communicate with Bob. If Charlie searches Alices computer and he finds Bobs public key, then he will know that Alice has been or intended to communicate with Bob. To counter this, Alice can randomize Bobs public key and store only the randomized key. The randomized key will be functionally equivalent to the public key associated with Bob, with the exception that nobody may associate it with Bob.
